---
title: Proposal for Project
subtitle: Milestone 3 - Billionaires 
authors: Garret Smith, Daniel Gallagher, Parishi Patel, Ava Byrd
---

```{r}
#| label: load-pkgs
#| message: false
library(tidyverse)
library(dplyr)

```

# Data 1

## Introduction and data

-   Identify the source of the data.

    Our data is sourced from the Peterson Institute for International Economics, and is presented by Nonresident Senior Fellow Caroline Freund and former research analyst Sarah Oliver.
    However, the original data from is from the Forbes World’s Billionaires lists from 1996-2014.

-   State when and how it was originally collected (by the original data curator, not necessarily how you found the data).

    -   The original data was collected by the Forbes World’s Billionaires lists from 1996-2014, however, scholars from the Peterson Institute for International economics compiled it into a dataset that was published in 2016, and added more variables such as the source of wealth, namely, "self made" or "inherited."

-   Write a brief description of the observations.

The dataset provides a comprehensive profile of billionaires, capturing key attributes like personal, company, location, and wealth information.

Personal Details: Includes the billionaire’s name, rank (showing relative wealth), age, and gender.
Age may be missing (marked as -1).
Company Information: Lists the company name, founding year, the billionaire’s relationship to the company (like “founder”), business sector, and type.
Location: Details the billionaire’s citizenship, country code, country’s GDP, and region.
Wealth: Describes their net worth in billions, the origin of their wealth (such as emerging markets, inheritance), industry, and the nature of their involvement (founder or political).
These fields provide insight into the billionaire’s background, the economic context of their country, and the nature of their wealth accumulation.

We plan on using all 1996, 2001 and 2014 data.
Rather than using data from just one year, using data from all of these years allows us to draw associations and patterns across the different years.
For this reason, it is okay for billionares to repeat.

Visualization 1: Gender vs Average Wealth

```{r}
library(ggplot2)
billionaires <- read.csv("data/billionaires.csv")

billionaires |>
  group_by(demographics.gender) |>
  summarize(mean_networth = mean(wealth.worth.in.billions))

billionaires_averages <-
billionaires |>
  group_by(demographics.gender) |>
  summarize(mean_networth = mean(wealth.worth.in.billions))

ggplot(billionaires_averages, aes(x = demographics.gender , y = mean_networth)) +
  geom_col() +
  labs(title = "Billionaire Wealth Types", x = "Demographics.Gender", y = "Wealth.Worth.In.Billions") +
  theme_minimal()




```

Visualization 2: Industry vs Avg.
Networth

```{r}

industry_averages <-
billionaires |>
  group_by(wealth.how.industry) |>
  summarize(mean_networth = mean(wealth.worth.in.billions))

ggplot(industry_averages, aes(x = wealth.how.industry , y = mean_networth)) +
  geom_col() +
  labs(title = "Industry vs Average Networth", x = "Industry", y = "Wealth.Worth.In.Billions") +
  theme_minimal()
 scale_x_discrete(expand = expansion(mult = c(0.1, 0.1)))
```

-   

<!-- -->

-   Address ethical concerns about the data, if any.

    Some ethical concerns about the data can be the privacy violations of the billionaires.
    Although the information is available to the public, the dataset could draw further attention to the billionaires which can lay the foundation for negative sentiment towards them.
    Additionally, as Freund and Oliver mention in their analysis, there is potential for several billionaires to be left off of the list due to a lack of reporting due to privacy concerns.
    This implies that there is potential that our research and outcomes are not representative of the global population of billionaires.

## Research question

-   A well formulated research question. (You may include more than one research question if you want to receive feedback on different ideas for your project. However, one per data set is required.)

How does the type of billionaire and the industry of the billionaire affect the billionaire’s net worth?

Revised Question: Across 1996, 2001, and 2014, how did the wealth origin, industry, gender and region affect the net worth of billionaires? 

-   Statement on why this question is important.

This is important because it can investigate the likelihood of an individual becoming a billionaire if they are not inheriting wealth, as many billionaires on the list have inherited wealth.
Some individuals aspire to be multi-billionaires, and this question investigates how realistic it is for someone to aspire for such wealth status.
Additionally, it provides insight into which industries are more likely for billionaires to cross the billionaire threshold.
This question also addresses the likelihood of women achieving a certain status within billionaires while also displaying that women make up a small percentage of billionaires.
Additionally, identifying the region of origin can correlate with industry type, which can provide valuable information into smaller relationships within the dataset.
Lastly, the question addresses the changes within the billionaire status as time goes on.

A description of the research topic along with a concise statement of your hypotheses on this topic.

-   This research explores the relationship between a billionaire’s type (whether they are self-made, have inherited wealth, or politically influenced wealth) and the industry in which they made their fortune, with a focus on how these factors impact their net worth. This research topic also examines various categories of billionaires, such as founders in technology, finance, and heirs in traditional industries. Overall, this research topic aims to discover patterns in the accumulation of wealth by individuals in these industries. The findings may highlight how certain industries or billionaire types yield higher net worth and give insight as to how wealth is accumulated.

We hypothesize that: The likelihood of becoming a billionaire is influenced by an individual's age, gender, and geographic region, with certain patterns emerging due to societal and economic factors.

-   Identify the types of variables in your research question. Categorical? Quantitative?

Billionaire Type: Categorical

Billionaire's Net Worth: Quantitative

Billionaire's Industry: Categorical

Billionaire's Gender: categorical

Billionaire's Region: categorical

## Glimpse of data

```{r}
#| label: load-data-1

billionaires <- read_csv("data/billionaires.csv")

glimpse(billionaires)

## Condensing sectors 

data <- billionaires |>
  mutate(
    company.sector = str_trim(tolower(
      company.sector
      )
    )
  ) |>  
  # Normalize company.sector values
  mutate(broad_industry = case_when(
    wealth.how.industry %in% c(
      "Technology-Computer", "Technology-Medical", "Software", 
      "IT Services", "Hardware", "Semiconductors"
      ) ~ "Technology",
    wealth.how.industry %in% c(
      "Banking", "Investment", "Private Equity", 
      "Hedge Funds", "Money Management", "Diversified financial", 
      "Private equity/leveraged buyout", "banking", "Venture Capital"
      ) ~ "Finance",
    wealth.how.industry %in% c(
      "Automotive", "Aerospace", "Industrial", 
      "Non-consumer industrial"
      ) ~ "Manufacturing",
    wealth.how.industry %in% c(
      "E-commerce", "Retail", "Fashion", "Retail, Restaurant"
      ) ~ "Retail",
    wealth.how.industry %in% c(
      "Pharmaceuticals", "Healthcare Services", "Biotech"
      ) ~ "Healthcare",
    wealth.how.industry %in% c(
      "Real Estate", "Construction", "Constrution"
      ) ~ "Real Estate",
    wealth.how.industry %in% c(
      "Media", "Entertainment"
      ) ~ "Media & Entertainment",
    wealth.how.industry %in% c(
      "Food Processing", "Beverage"
      ) ~ "Food & Beverage",
    wealth.how.industry %in% c(
      "Agriculture", "Energy", "Mining", "Mining and metals"
      ) ~ "Natural Resources",
    wealth.how.industry == "Logistics" ~ "Logistics",
    # Temporary NA for ungrouped values
    wealth.how.industry %in% c("Other", "0", "services") ~ NA_character_,  
    TRUE ~ wealth.how.industry
    )
  ) |>
  mutate(broad_industry = case_when(
    is.na(broad_industry) & company.sector %in% c(
      "technology", "it services", "software", "internet") ~ "Technology",
    is.na(broad_industry) & company.sector %in% c(
      "finance", "banking", "investment", "hedge funds", "financial services"
      ) ~ "Finance",
    is.na(broad_industry) & company.sector %in% c(
      "automotive", "manufacturing", "aerospace", "industrial"
      ) ~ "Manufacturing",
    is.na(broad_industry) & company.sector %in% c(
      "e-commerce", "retail", "fashion", "luxury goods"
      ) ~ "Retail",
    is.na(broad_industry) & company.sector %in% c(
      "pharmaceuticals", "healthcare", "biotech", "medical"
      ) ~ "Healthcare",
    is.na(broad_industry) & company.sector %in% c(
      "real estate", "construction"
      ) ~ "Real Estate",
    is.na(broad_industry) & company.sector %in% c(
      "media", "entertainment", "publishing"
      ) ~ "Media & Entertainment",
    is.na(broad_industry) & company.sector %in% c(
      "food processing", "beverage", "groceries", "food"
      ) ~ "Food & Beverage",
    is.na(broad_industry) & company.sector %in% c(
      "agriculture", "energy", "mining", "oil", "natural resources"
      ) ~ "Natural Resources",
    # Retain any pre-assigned broad_industry values
    is.na(broad_industry) & company.sector == "logistics" ~ "Logistics",
    TRUE ~ broad_industry  
  ))

billionaires |>
  count(wealth.how.industry)

data |>
  count(broad_industry)

```

# Data 2

## Introduction and data

-   Identify the source of the data.

The Washington Post

-   State when and how it was originally collected (by the original data curator, not necessarily how you found the data).

In 2015, the Washington Post began collecting data of officer involved shootings including many variables about the encounters and suspects.
The Post manually went through multiple sources including local news reports, information from law enforcement websites and social media, and from other databases such as Fatal Encounters and the Killed by Police project.

-   Write a brief description of the observations.

Each observation is a suspect that was shot by police and includes the date of incident, gender, race, threat level, location, whether a body camera was present, whether the suspect was fleeing, and the mental state.

-   Address ethical concerns about the data, if any.

The names of the victims are included in the data which could be a potential invasion of privacy, especially if they were found to be unjustly shot.
There is also an accompanying dataset that includes the police/sheriff offices that were involved, which could raise concerns if the shooting was justified and the officers are attacked online by the public.

## Research question

-   **Research Question Ideas**

1.  How does the race of a victim influence the perceived threat level in police encounters across different U.S. states?
2.  How does the use of body cameras affect the number of shootings of higher and lower threat suspects across different regions.

NOTE (11/12): The perceived threat level variable isn't the best response variable because it only has values "attack", "other", and "undetermined".
We will not use this research question and will most likely use data set 1.
Here is a alternative research question for this dataset.

2.  How does the use of body cameras affect the number of shootings of higher and lower threat suspects across different regions.

NOTE(11/12): This research question uses the number of shootings in each region as a response variable instead of the threat level.

-   Statement on why this question is important.

1.  Understanding the influence of race on perceived threat levels in police encounters is important for address systemic bias in the use of force. We can identify patterns of racial bias by examining how these perceptions vary across states and help inform policies aimed at reducing racially motivated police violence.
2.  This question seeks to understand whether body cameras can reduce shootings and whether the region plays a role in that statistic. This question also looks at whether the suspect was armed because this is an important metric in understanding whether the shooting was justified or not. While there is a lot of context missing from all these interactions, knowing whether they were armed with an actual gun or potentially a fake one could give better insight into a perceived threat. This threat level can also be carefully considered against the actual threat level variable to see better understand if one or another is a better objective representative of an unjustified or justified shooting.

-   **Description**

1.  We will explore how racial perceptions shape the perceived threat level of individuals in police encounters, with a focus on differences across states. The hypothesis is that individuals from minority racial groups, particularly Black and Hispanic individuals, are more likely to be perceived as threatening by law enforcement compared to White individuals, and that these perceptions may vary based on the state’s demographic and political environment.
2.  Shootings are an important political and social issue in this country so a question that uses this police shooting data to understand if suspects are shot less when body cameras are present in low threat level scenarios. The thought would be that by having a body camera might reduce the shootings that occur due to an added level of accountability. While we cannot test given all interactions with police when people weren’t shot, by comparing body-camera shootings to no body-camera shootings, we can potentially see the varying proportions of low threat level shooting depending on whether there is that oversight. This can also be applied regionally to see if factors that exist geographically to create this trend.

-   **Variable Types**

1.  

    a.  Race of the victim: Categorical (Nominal)

    b.  Perceived threat level: Categorical (Nominal)

    c.  State: Categorical (Nominal)

2.  

    a.  Body Camera: Categorical (Logical)

    *Threat*

    b.  Arms Category: Categorical (Nominal)
    c.  Threat Level: Categorical (Nominal)

    *Location*

    d.  Region: Categorical (Nominal)
    e.  Longitude: Numerical (Continuous)
    f.  Latitude: Numerical (Continuous)

3.  

    1.  Body Camera: Logical

    2.  Arms Category: Categorical (Nominal)

    3.  Region: Categorical (Nominal)

    4.  Longitude: Numerical (Continuous)

    5.  Latitude: Numerical (Continuous)

## Glimpse of data